---
title: "Advanced Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## Overview

This vignette covers advanced features of **lwdid**: multiple estimators,
sensitivity analysis, diagnostics, wild cluster bootstrap, randomization
inference, data preprocessing, parallel computation, and result export.

```{r load}
library(lwdid)
data(smoking)
data(castle)
```

## Multiple Estimators

lwdid supports four estimators, all applied to the transformed
cross-sectional data:

| Estimator | `estimator =` | Controls Required | Double Robust |
|-----------|--------------|-------------------|---------------|
| Regression Adjustment | `"ra"` | Optional | No |
| Inverse Probability Weighting | `"ipw"` | Yes | No |
| Doubly Robust | `"ipwra"` | Yes | Yes |
| Propensity Score Matching | `"psm"` | Yes | No |

Use `controls` for RA and IPWRA outcome-regression controls. Use
`ps_controls` for the IPW, IPWRA, and PSM propensity-score model. When
`ps_controls` is omitted, the propensity-score estimators fall back to
`controls`.

### IPW

```{r ipw, eval=FALSE}
result_ipw <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "gvar", rolling = "demean", estimator = "ipw",
  ps_controls = c("police", "income", "poverty"),
  aggregate = "cohort", control_group = "never_treated"
)
```

### IPWRA (Doubly Robust)

```{r ipwra, eval=FALSE}
result_ipwra <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "gvar", rolling = "demean", estimator = "ipwra",
  controls = c("police", "income", "poverty"),
  ps_controls = c("police", "income", "poverty"),
  aggregate = "cohort", control_group = "never_treated"
)
```

### PSM

```{r psm, eval=FALSE}
result_psm <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "gvar", rolling = "demean", estimator = "psm",
  ps_controls = c("police", "income", "poverty"),
  aggregate = "cohort", control_group = "never_treated"
)
```

## Transformation Comparison

```{r transforms}
res_dm <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", rolling = "demean"
)

res_dt <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", rolling = "detrend"
)

cat("Demean ATT:", round(res_dm$att, 4), "\n")
cat("Detrend ATT:", round(res_dt$att, 4), "\n")
```

For quarterly data, use `"demeanq"` or `"detrendq"` with the
`season_var` parameter.

## Variance-Covariance Estimation

| `vce` | Description |
|-------|-------------|
| `NULL` | OLS (homoskedastic) |
| `"robust"` / `"hc1"` | HC1 heteroskedasticity-robust |
| `"hc0"` -- `"hc4"` | HC-family (HC3 recommended for small $N$) |
| `"cluster"` | Cluster-robust (requires `cluster_var`) |

```{r vce-example}
# HC1 for smoking data (single treated unit makes HC3 unstable)
result_hc1 <- suppressWarnings(lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", vce = "hc1"
))
summary(result_hc1)
```

## Comparing Multiple Specifications

The `compare()` function provides a structured way to compare results from
multiple specifications (different estimators, transformations, or control
variables). This is essential for robustness analysis in the Lee & Wooldridge
(2025) framework, where four estimators exhibit different consistency and
efficiency properties under varying assumptions:

- **RA** (Regression Adjustment): BLUE under correct outcome model
- **IPW** (Inverse Probability Weighting): consistent when propensity score
  model is correct
- **IPWRA** (Doubly Robust): consistent if either model is correct
- **PSM** (Propensity Score Matching): nonparametric kernel matching

If all specifications yield similar point estimates, results are robust to
model specification.

### Basic Usage: Comparing Estimators

```{r compare-basic, eval=FALSE}
# Estimate with different estimators
res_ra <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                gvar = "gvar", rolling = "demean", estimator = "ra",
                aggregate = "overall")
res_ipw <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                 gvar = "gvar", rolling = "demean", estimator = "ipw",
                 ps_controls = c("police", "income", "poverty"),
                 aggregate = "overall")
res_ipwra <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                   gvar = "gvar", rolling = "demean", estimator = "ipwra",
                   controls = c("police", "income", "poverty"),
                   ps_controls = c("police", "income", "poverty"),
                   aggregate = "overall")

# Compare specifications side by side
# Output: table with ATT, SE, CI, significance stars, model metadata
compare(RA = res_ra, IPW = res_ipw, IPWRA = res_ipwra)
```

### Advanced Usage: Comparing Transformations

```{r compare-transform, eval=FALSE}
# Demean vs Detrend comparison
res_demean <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                    gvar = "gvar", rolling = "demean", estimator = "ra",
                    aggregate = "overall")
res_detrend <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                     gvar = "gvar", rolling = "detrend", estimator = "ra",
                     aggregate = "overall")

# Similar estimates suggest robustness to trend specification
compare(Demean = res_demean, Detrend = res_detrend)
```

### Type Parameter: Overall vs Effects

The `type` argument controls how comparison results are reported:

- `type = "overall"` (default): compares the scalar ATT from each specification
- `type = "effects"`: compares cohort-level or period-level effects

```{r compare-effects, eval=FALSE}
# Effects-level comparison shows cohort-by-cohort differences
compare(RA = res_ra, IPWRA = res_ipwra, type = "effects")

# Output columns include:
# - Model name, ATT estimate, standard error
# - 95% confidence interval [ci_lower, ci_upper]
# - Significance stars (*** p<0.01, ** p<0.05, * p<0.1)
# - Model metadata: estimator, transformation, N observations
```

## Sensitivity Analysis

```{r sensitivity}
sens <- suppressWarnings(lwdid_sensitivity(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", type = "all", verbose = FALSE
))
print(sens)
```

### Pre-Period Sensitivity

```{r pre-period-sens, eval=FALSE}
pre_sens <- lwdid_sensitivity_pre_period(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post"
)
```

## Diagnostics

### Parallel Trends Testing

```{r pt-test}
pt <- lwdid_test_parallel_trends(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "gvar"
)
print(pt)
```

### Transformation Recommendation

```{r recommend}
rec <- lwdid_recommend_transformation(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "gvar"
)
print(rec)
```

### Full Diagnostic Suite

```{r diagnose}
diag <- suppressWarnings(lwdid_diagnose(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "gvar"
))
summary(diag)
```

## Data Preprocessing

`aggregate_to_panel()` collapses micro-level data to panel format:

```{r aggregate, eval=FALSE}
agg_result <- aggregate_to_panel(
  data = survey_data,
  unit_var = "state",
  time_var = "year",
  outcome_var = "income",
  weight_var = "survey_weight",
  treatment_var = "treated",
  min_cell_size = 30
)
panel <- as.data.frame(agg_result)
```

## Wild Cluster Bootstrap

For inference with few clusters, use `vce = "bootstrap"` with a
`cluster_var`:

```{r wcb, eval=FALSE}
result_wcb <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", rolling = "demean",
  vce = "bootstrap", cluster_var = "state",
  wcb_reps = 999, wcb_seed = 42,
  use_fwildclusterboot = FALSE
)
summary(result_wcb)
```

WCB output reports both requested and actual bootstrap draws. With
Rademacher weights and at most 12 clusters, `lwdid` uses full sign-pattern
enumeration, so the actual draw count can differ from `wcb_reps`.

When `vce = "cluster"` is specified and the number of clusters is below
20, lwdid automatically upgrades to the Wild Cluster Bootstrap
(`auto_wcb = TRUE` by default).

## Randomization Inference

Randomization-based p-values without asymptotic normality:

```{r ri, eval=FALSE}
result_ri <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", ri = TRUE, ri_method = "permutation",
  rireps = 500, seed = 42
)
cat("RI p-value:", result_ri$ri_pvalue, "\n")
```

RI output includes `ri_observed_stat`, the observed statistic used for the
randomization tail comparison. In common-timing designs it is the top-level ATT.
In staggered designs, `ri_target` records the target: `aggregate = "overall"`
uses the overall ATT, `aggregate = "cohort"` uses the selected cohort ATT, and
`aggregate = "none"` uses the first finite cohort-time ATT.

## Parallel Computation

Enable parallel processing for staggered designs:

```{r parallel, eval=FALSE}
result_par <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "gvar", rolling = "demean", aggregate = "cohort",
  control_group = "never_treated", parallel = TRUE, n_cores = 4
)
```

Or configure the parallel backend explicitly:

```{r parallel-setup, eval=FALSE}
setup_parallel(n_workers = 4, strategy = "multisession")
# All subsequent lwdid() calls use parallel backend
result <- lwdid(...)
# Reset when done
setup_parallel(strategy = "sequential")
```

## Integration with Broom and modelsummary

lwdid provides `tidy()` and `glance()` S3 methods (compatible with the
**broom** package interface), as well as fixest-style accessor functions.
These enable seamless integration with **modelsummary** and other tidyverse
workflows.

### tidy(): Tidy Coefficient Tables

The `tidy()` method extracts results in a standardized data.frame format.
The `type` argument selects which estimates to report:

```{r tidy-example, eval=FALSE}
library(lwdid)
result <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                gvar = "gvar", rolling = "demean", estimator = "ra",
                aggregate = "overall")

# Overall ATT (default)
# Returns: data.frame with columns term, estimate, std.error, statistic,
#          p.value, conf.low, conf.high
tidy(result)

# Cohort-level effects
# Returns one row per cohort with cohort-specific ATT
tidy(result, type = "effects")

# Event-time aggregation
# Returns one row per relative event-time period
tidy(result, type = "event_time")
```

### glance(): Model-Level Statistics

`glance()` returns a one-row data.frame of model-level summary statistics:

```{r glance-example, eval=FALSE}
# Returns: nobs, n_treated, n_control, n_cohorts, n_periods,
#          estimator, transformation, vce_type, r.squared (if applicable)
glance(result)
```

### Fixest-Style Accessors

For users familiar with the **fixest** package interface:

```{r fixest-style, eval=FALSE}
# Standard errors
se(result)

# t-statistics
tstat(result)

# p-values
pvalue(result)

# Full coefficient table (estimate, se, t-stat, p-value)
coeftable(result)
```

### modelsummary Integration

With **modelsummary** installed, lwdid results can be directly passed to
`modelsummary()` for publication-quality tables:

```{r modelsummary-example, eval=FALSE}
library(modelsummary)

# Compare multiple specifications in one table
res_ra <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                gvar = "gvar", rolling = "demean", estimator = "ra",
                aggregate = "overall")
res_ipwra <- lwdid(data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
                   gvar = "gvar", rolling = "demean", estimator = "ipwra",
                   controls = c("police", "income", "poverty"),
                   ps_controls = c("police", "income", "poverty"),
                   aggregate = "overall")

# Produces a formatted comparison table with estimates, SEs, and fit stats
modelsummary(list(RA = res_ra, IPWRA = res_ipwra))

# Export to LaTeX or Word
modelsummary(list(RA = res_ra, IPWRA = res_ipwra),
             output = "results.tex")
modelsummary(list(RA = res_ra, IPWRA = res_ipwra),
             output = "results.docx")
```

## Exporting Results

```{r export, eval=FALSE}
to_latex(res_dm, file = "results_table.tex")
to_csv(res_dm, file = "results.csv")
```

## References

Lee, S. and Wooldridge, J.M. (2026). "A Simple Transformation Approach to
Difference-in-Differences Estimation for Panel Data."

Lee, S., Wooldridge, J.M. and Hur, E.K. (2026). "Rolling Difference-in-Differences
Estimation for Small and Large Panels."
