---
title: "Advanced Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## Overview

This vignette covers advanced features of **lwdid**: multiple estimators,
sensitivity analysis, diagnostics, wild cluster bootstrap, randomization
inference, data preprocessing, parallel computation, and result export.

```{r load}
library(lwdid)
data(smoking)
data(castle)
```

## Multiple Estimators

lwdid supports four estimators, all applied to the transformed
cross-sectional data:

| Estimator | `estimator =` | Controls Required | Double Robust |
|-----------|--------------|-------------------|---------------|
| Regression Adjustment | `"ra"` | Optional | No |
| Inverse Probability Weighting | `"ipw"` | Yes | No |
| Doubly Robust | `"ipwra"` | Yes | Yes |
| Propensity Score Matching | `"psm"` | Yes | No |

### IPW

```{r ipw, eval=FALSE}
result_ipw <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "effyear", rolling = "demean", estimator = "ipw",
  controls = c("police", "income", "poverty"),
  aggregate = "cohort"
)
```

### IPWRA (Doubly Robust)

```{r ipwra, eval=FALSE}
result_ipwra <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "effyear", rolling = "demean", estimator = "ipwra",
  controls = c("police", "income", "poverty"),
  aggregate = "cohort"
)
```

### PSM

```{r psm, eval=FALSE}
result_psm <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "effyear", rolling = "demean", estimator = "psm",
  controls = c("police", "income", "poverty"),
  aggregate = "cohort"
)
```

## Transformation Comparison

```{r transforms}
res_dm <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", rolling = "demean"
)

res_dt <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", rolling = "detrend"
)

cat("Demean ATT:", round(res_dm$att, 4), "\n")
cat("Detrend ATT:", round(res_dt$att, 4), "\n")
```

For quarterly data, use `"demeanq"` or `"detrendq"` with the
`season_var` parameter.

## Variance-Covariance Estimation

| `vce` | Description |
|-------|-------------|
| `NULL` | OLS (homoskedastic) |
| `"robust"` / `"hc1"` | HC1 heteroskedasticity-robust |
| `"hc0"` -- `"hc4"` | HC-family (HC3 recommended for small $N$) |
| `"cluster"` | Cluster-robust (requires `cluster_var`) |

```{r vce-example}
# HC1 for smoking data (single treated unit makes HC3 unstable)
result_hc1 <- suppressWarnings(lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", vce = "hc1"
))
summary(result_hc1)
```

## Sensitivity Analysis

```{r sensitivity, eval=FALSE}
sens <- lwdid_sensitivity(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", type = "all", verbose = FALSE
)
print(sens)
```

### Pre-Period Sensitivity

```{r pre-period-sens, eval=FALSE}
pre_sens <- lwdid_sensitivity_pre_period(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post"
)
```

## Diagnostics

### Parallel Trends Testing

```{r pt-test, eval=FALSE}
pt <- lwdid_test_parallel_trends(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "effyear"
)
print(pt)
```

### Transformation Recommendation

```{r recommend, eval=FALSE}
rec <- lwdid_recommend_transformation(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "effyear"
)
print(rec)
```

### Full Diagnostic Suite

```{r diagnose, eval=FALSE}
diag <- lwdid_diagnose(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "effyear"
)
summary(diag)
```

## Data Preprocessing

`aggregate_to_panel()` collapses micro-level data to panel format:

```{r aggregate, eval=FALSE}
agg_result <- aggregate_to_panel(
  data = survey_data,
  unit_var = "state",
  time_var = "year",
  outcome_var = "income",
  weight_var = "survey_weight",
  treatment_var = "treated",
  min_cell_size = 30
)
panel <- as.data.frame(agg_result)
```

## Wild Cluster Bootstrap

For inference with few clusters:

```{r wcb, eval=FALSE}
wcb <- wild_cluster_bootstrap(
  data = castle, y = "lhomicide", ivar = "sid",
  tvar = "year", gvar = "effyear",
  cluster = "sid", B = 999
)
print(wcb)
```

## Randomization Inference

Exact p-values without distributional assumptions:

```{r ri, eval=FALSE}
result_ri <- lwdid(
  data = smoking, y = "lcigsale", ivar = "state", tvar = "year",
  d = "d", post = "post", ri = TRUE, rireps = 500
)
cat("RI p-value:", result_ri$ri_pvalue, "\n")
```

## Parallel Computation

Enable parallel processing for staggered designs:

```{r parallel, eval=FALSE}
result_par <- lwdid(
  data = castle, y = "lhomicide", ivar = "sid", tvar = "year",
  gvar = "effyear", rolling = "demean", aggregate = "cohort",
  parallel = TRUE, n_cores = 4
)
```

Or configure the parallel backend explicitly:

```{r parallel-setup, eval=FALSE}
setup_parallel(n_workers = 4, strategy = "multisession")
# All subsequent lwdid() calls use parallel backend
result <- lwdid(...)
# Reset when done
setup_parallel(strategy = "sequential")
```

## Exporting Results

```{r export, eval=FALSE}
to_latex(result, file = "results_table.tex")
to_csv(result, file = "results.csv")
```

## References

Lee, S. and Wooldridge, J.M. (2025). "A Simple Transformation Approach to
Difference-in-Differences Estimation for Panel Data."

Lee, S. and Wooldridge, J.M. (2026). "Simple Approaches to Inference with
Difference-in-Differences Estimators with Small Cross-Sectional Sample
Sizes."
