---
title: "Getting Started with lwdid"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with lwdid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

## Introduction

The **lwdid** package implements the Lee and Wooldridge (2025, 2026) simple
transformation approach for Difference-in-Differences (DiD) estimation with
panel data. The key innovation is a unit-level time-series transformation
that converts panel DiD into a cross-sectional treatment effects problem.

**Key features:**

- **Two designs**: Common Timing and Staggered Adoption
- **Four estimators**: RA, IPW, IPWRA, PSM
- **Robust inference**: HC0--HC4, cluster-robust, Wild Cluster Bootstrap,
  Fisher Randomization Inference
- **Exact small-sample inference**: valid even with $N_1 = 1$
- **Diagnostics**: parallel trends tests, sensitivity analysis, clustering
  diagnostics

## Installation

```{r eval=FALSE}
# From GitHub
# install.packages("devtools")
devtools::install_github("gorgeousfish/lwdid-r")
```

## Data Format

lwdid requires **long-format panel data** where each row is one
unit-period observation. Required columns:

| Role | Parameter | Type |
|------|-----------|------|
| Unit ID | `ivar` | integer or character |
| Time | `tvar` | integer |
| Outcome | `y` | numeric |
| Treatment (Common Timing) | `d` + `post` | binary (0/1) |
| Treatment (Staggered) | `gvar` | integer or NA |

## Demean vs Detrend

The `rolling` parameter controls pre-treatment transformation:

- **`"demean"`** (default): Subtracts the pre-treatment mean.
  Requires parallel trends. At least 1 pre-treatment period.
- **`"detrend"`**: Removes unit-specific linear trends. Allows
  heterogeneous trends. At least 2 pre-treatment periods.

## Common Timing: California Proposition 99

The `smoking` dataset contains the classic California tobacco control
study (1970--2000). California is the sole treated state ($N_1 = 1$),
with treatment starting in 1989.

```{r common-timing-basic}
library(lwdid)
data(smoking)

result <- lwdid(
  data = smoking,
  y = "lcigsale",
  ivar = "state",
  tvar = "year",
  d = "d",
  post = "post",
  rolling = "demean",
  estimator = "ra"
)

print(result)
```

### Detrending

When unit-specific trends are suspected:

```{r common-timing-detrend}
result_dt <- lwdid(
  data = smoking,
  y = "lcigsale",
  ivar = "state",
  tvar = "year",
  d = "d",
  post = "post",
  rolling = "detrend",
  estimator = "ra"
)

summary(result_dt)
```

## Staggered Adoption: Castle Doctrine Laws

The `castle` dataset examines the effect of Castle Doctrine laws on
homicide rates across US states. States adopted the law at different
times between 2005 and 2009.

```{r staggered-basic}
data(castle)

result_stag <- lwdid(
  data = castle,
  y = "lhomicide",
  ivar = "sid",
  tvar = "year",
  gvar = "gvar",
  rolling = "demean",
  estimator = "ra",
  aggregate = "cohort"
)

print(result_stag)
```

### Event Study

```{r staggered-event-study, eval=requireNamespace("ggplot2", quietly=TRUE)}
result_event <- lwdid(
  data = castle,
  y = "lhomicide",
  ivar = "sid",
  tvar = "year",
  gvar = "gvar",
  rolling = "demean",
  estimator = "ra",
  aggregate = "event_time"
)

plot(result_event)
```

## Control Variables

Controls improve efficiency and relax the parallel trends assumption.
They must be **time-invariant** (or pre-treatment averages):

```{r controls, eval=FALSE}
result_ctrl <- lwdid(
  data = castle,
  y = "lhomicide",
  ivar = "sid",
  tvar = "year",
  gvar = "gvar",
  rolling = "demean",
  estimator = "ra",
  controls = c("police", "income", "poverty", "unemployrt"),
  aggregate = "cohort"
)
```

## Robust Variance Estimation

| `vce` | Description |
|-------|-------------|
| `NULL` | OLS (homoskedastic) |
| `"robust"` / `"hc1"` | HC1 heteroskedasticity-robust |
| `"hc3"` | HC3 (recommended for small samples) |
| `"cluster"` | Cluster-robust (requires `cluster_var`) |

```{r vce}
# HC1 is recommended for the smoking data (single treated unit makes HC3
# numerically unstable due to hat value = 1)
result_hc1 <- suppressWarnings(lwdid(
  data = smoking,
  y = "lcigsale",
  ivar = "state",
  tvar = "year",
  d = "d",
  post = "post",
  rolling = "demean",
  estimator = "ra",
  vce = "hc1"
))

summary(result_hc1)
```

## Extracting Results

```{r extract}
coef(result)
confint(result)
vcov(result)
nobs(result)
```

## Parallel Computation

For staggered designs with many cohort-period pairs, enable parallel
processing:

```{r parallel, eval=FALSE}
result_par <- lwdid(
  data = castle,
  y = "lhomicide",
  ivar = "sid",
  tvar = "year",
  gvar = "gvar",
  rolling = "demean",
  estimator = "ra",
  aggregate = "cohort",
  parallel = TRUE,
  n_cores = 4
)
```

## Next Steps

- `vignette("advanced-usage")`: IPW, IPWRA, PSM, sensitivity analysis,
  diagnostics, WCB, RI, data preprocessing
- `vignette("methodology")`: mathematical framework and identification

## References

Lee, S. and Wooldridge, J.M. (2025). "A Simple Transformation Approach to
Difference-in-Differences Estimation for Panel Data."

Lee, S. and Wooldridge, J.M. (2026). "Simple Approaches to Inference with
Difference-in-Differences Estimators with Small Cross-Sectional Sample
Sizes."
